Designing Modular Software: A Case Study in Introductory Statistics

### 2016-09-30 jcgs-paper-2016/paper.Rmd
add table

Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation and helps abstract away some of the more challenging programming aspects of working directly with R. Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application. Several recent software packages have built up Shiny to provide a web-based system based on R. One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge. Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting. Another package is called Radiant [@radiant]

>>>>

Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation and helps abstract away some of the more challenging programming aspects of working directly with R. Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application. Several recent software packages have built up Shiny to provide a web-based system based on R. One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge. Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting. Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis. While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student. An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}. Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default. For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.

```{r, echo=FALSE}
software_compare <- data.frame("Software package" = "intRo",
                               "GUI-based" = "Yes",
                               "Installation required" = "No",
                               "Modular" = "Yes",
                               "Web-based" = "Yes",
                               "Extensible" = "Yes",
                               "Fosters Reproducibility" = "Yes",
                               "Open Source" = "Yes",
                               "Functionality" = "Limited",
                               stringsAsFactors = FALSE)

software_compare <- rbind(software_compare,
                          c("JMP", "Yes", "Yes", "Partial", "No", "Partial", "No", "No", "Full"),
                          c("R", "Partial", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "Full"),
                          c("R Commander", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
                          c("Deducer", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
                          c("MOSAIC", "No", "Yes", "No", "No", "No", "Yes", "Yes", "Limited"),
                          c("iNZight Lite", "Yes", "No", "Yes", "Yes", "Partial", "No", "Yes", "Limited"),
                          c("Radiant", "Yes", "Yes", "Yes", "Yes", "Partial", "Yes", "Yes", "Moderate"))

software_compare <- software_compare[, -8]

knitr::kable(software_compare, 
             caption="\\label{tab:compare}A comparison of statistical software packages across the metrics of usability, modularity, extensibility, and reproducibility. Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.", 
             col.names = c("Software", "GUI", "Install", "Modular", "Web-based", "Extensible", "Reproducible", "Functionality"))
```
